Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-12-02 edubots_surveys.Rnw
Latest changes

Moreover, this study clarifies the differences between the teachers’
perspectives and/or intentions and that of their students when using the
messaging applications in higher education.  Teachers tend to adopt technologies
supported by their institutions.  This could be to ensure that their
universities can oversee their efforts in supporting their students during the
learning process.  Another reason could be their familiarity with the technology
provided by their institutions. On the other hand, students use the
non-institution messaging applications to form informal discussion groups with
their peers.  It is worth noting that peer support and collaboration is
inseparable from learning \cite{timmis2012constant} and correlates positively
with higher retention in higher education \cite{o2014front}.  Therefore, both
perspectives are complementary and play different roles in promoting the
learning process.

@@

% What kind of chatbots are preferred? (suggested by Fran)
Once that initial introduction is done, teachers (and students as well) will
probably prefer the kind of bots that alleviate bureaucratic or repetitive
tasks, such as answering frequent questions or answers on class or assignment
schedules. These will help the introduction of more complex chatbots that will
affect more directly learning outcomes, such as chatbots that help students
integrate in the class, or are able to identify (and address) learning problems
in students or in groups of them. These as well should be accompanied by
analytics on student interaction, as well as possibly some natural language
processing (in vernacular language) that will help assess the general mood of
the class, and how different material (or external factors) affect it.

@@

\begin{figure}[h!tbp]
\begin{center}
\includegraphics[width=0.85\textwidth]{figures/survey-1.png}
\end{center}
    \caption{\textcolor{blue}{Screen capture of the Spanish-language form used for the first survey}}
\label{figure:survey1}
\end{figure}
% Including questions in surveys
The {\bf first survey}, shown in its original form in Figure \ref{figure:survey1}, which \textcolor{blue}{focused on} the use of messaging apps in class, and which tried to address mainly RQ2 and RQ3, 
% Antonio - I guess it refers to RQ3. ;)
includes the following questions:

@@

\begin{figure}[h!tbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/survey-2.png}
\end{center}
    \caption{\textcolor{blue}{Screen capture of the Spanish-language form used for the survey}}
\label{figure:survey2}
\end{figure}
The {\bf second survey}, shown in its original form in Figure \ref{figure:survey2}, followed the same paths, only in this case there was a single questionnaire, and with a different scope, focusing on the kind of interaction media features that teachers value the most. %\textcolor{blue}{A capture of these questions is included in Figure \ref{fig:survey2}} % this sentence is repeated
% Olia - see my comment above.
It was filled by teachers that were also students in a Professional Development course (on the use of new technologies in higher education, around 1/4 of them) as well as University of Granada teachers who knew about it by emails from the authors, or from the vicedeanship for International Relations that included it in its newsletter. This means that there might be a higher proportion of 1) teachers with few years of experience 2) teachers from the computer science faculty and 3) teachers from Granada. We think, however that there's no explicit bias in this selection, although of course specific percentages will vary. These questions tried to address mainly RQ1 and RQ4.

>>>>

\textcolor{blue}{Comparing these results with those obtained in student surveys
in \cite{MoraChatbots2021:anon}, we have understood better} the differences between the teachers’
perspectives and/or intentions and that of their students when using the
messaging applications in higher education.  Teachers tend to adopt technologies
supported by their institutions.  This could be to ensure that their
universities can oversee their efforts in supporting their students during the
learning process.  Another reason could be their familiarity with the technology
provided by their institutions. On the other hand, students use the
non-institution messaging applications to form informal discussion groups with
their peers.  It is worth noting that peer support and collaboration is
inseparable from learning \cite{timmis2012constant} and correlates positively
with higher retention in higher education \cite{o2014front}.  Therefore, both
perspectives are complementary and play different roles in promoting the
learning process.

@@

\textcolor{blue}{What we are going to propose next is a possible process of
technology introduction that is compatible with the conclusions of this study,
but that would have to be piloted in order to check its value, and its
relationship with better learning outcomes, as well as higher teacher
satisfaction.}
Once that initial introduction \textcolor{blue}{of institutional messaging
automation tools} is done, teachers (and students as well) will
probably prefer the kind of bots that alleviate bureaucratic or repetitive
tasks, such as answering frequent questions or answers on class or assignment
schedules \textcolor{blue}{, as indicated by their answers to the respective
questions}. These will help the introduction of more complex chatbots that will
affect more directly learning outcomes, such as chatbots that help students
integrate in the class, or are able to identify (and address) learning problems
in students or in groups of them. These as well should be accompanied by
analytics on student interaction, as well as possibly some natural language
processing (in vernacular language) that will help assess the general mood of
the class, and how different material (or external factors) affect it.

@@

\begin{figure}[h!tbp]
\begin{center}
\includegraphics[width=0.85\textwidth]{figures/survey-1.png}
\end{center}
    \caption{\textcolor{blue}{Screen capture of the Spanish-language form used for the first survey}}
\label{figure:survey1}
\end{figure}
% Including questions in surveys
The {\bf first survey}, shown in its original form in Figure \ref{figure:survey1}, which \textcolor{blue}{focused on} the use of messaging apps in class, includes the following questions:

@@

\begin{figure}[h!tbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/survey-2.png}
\end{center}
    \caption{\textcolor{blue}{Screen capture of the Spanish-language form used for the survey}}
\label{figure:survey2}
\end{figure}
The {\bf second survey}, shown in its original form in Figure \ref{figure:survey2}, followed the same paths, only in this case there was a single questionnaire, and with a different scope, focusing on the kind of interaction media features that teachers value the most. %\textcolor{blue}{A capture of these questions is included in Figure \ref{fig:survey2}} % this sentence is repeated
% Olia - see my comment above.
It was filled by teachers that were also students in a Professional Development course (on the use of new technologies in higher education, around 1/4 of them) as well as University of Granada teachers who knew about it by emails from the authors, or from the vicedeanship for International Relations that included it in its newsletter. This means that there might be a higher proportion of 1) teachers with few years of experience 2) teachers from the computer science faculty and 3) teachers from Granada. We think, however that there's no explicit bias in this selection, although of course specific percentages will vary.
